---
title: "R Notebook"
output: html_notebook
---

# Day 1


##### Important packages
``` {r}
install.packages(c("car", "tidyverse", "mosaic", "openintro", "MASS"))
```


## Rstudio tour

Things to mention:
  Each pane:
    Notebook/script pane
    Console/terminal
    Env/history/connections/git
    Files/plots/packages/help
  
  Getting help (from command line or from pane)
  
  Projects (Organizing code)
    Make a project at this point

## Assigning, manipulating, re-assigning

``` {r}
# Canonical first thing you do with a new language
print("Hello World!")

x <- 1
x
X
x <- 10
x
x * 2
x
x <- x * 2
x
x <- x / 2
x
x <- "x"
x
x <- "x" + 2 # Errors
x <- "x" + "2" # Errors, works in python/js

a <- 1
b <- 2
c <- 3

a + b + c

a + b / c
(a + b) / c
d <- (a + b) / c

x # is unchanged

# Vectors
c(1,2,3)
test_vector <- c(TRUE, FALSE, NA)
test_vector[0]
test_vector[2]
test_vector[[2]]

# Lists
test_list <- list(1, 2,"hello", TRUE)
test_list[3]
test_list[[3]]

# Vector of factors
x <- factor(c("yes", "yes", "no", "yes", "no")) 
x

table(x)

unclass(x)

x <- factor(c("yes", "yes", "no", "yes", "no"),
             levels = c("yes", "no"))

x

unclass(x)


# Data frames
df <- data.frame(var1 = 1:4, var2 = c(T, T, F, F)) 
df
nrow(df)
ncol(df)

names(df)
names(df) <- c("id", "likes_ice_cream")
names(df)
df

df[1]
df[1, ]
df[, 2]
df$likes_ice_cream
df$id

df[[1]]
```


## Logical operators

``` {r}
1 == 1
1 == 0
1 != 0
NA == NA
NA == FALSE
NA == TRUE
NA == 0
NA == 1
is.na(NA)
1 == 1
!(1 == 1)

1 < 2
2 < 1
2 > 1
1 > 2

1 <= 2
2 >= 1

TRUE == !FALSE
!TRUE == FALSE
!TRUE == !FALSE

(1 == 1) && (2 == 2)
(1 == 2) && (2 == 2)
(1 == 1) && (2 == 1)
(1 == 2) && (2 == 1)

(1 == 1) || (2 == 2)
(1 == 2) || (2 == 2)
(1 == 1) || (2 == 1)
(1 == 2) || (2 == 1)

NA && is.na(NA)
NA || is.na(NA)

is.numeric(1)
is.numeric("hello")

is.character("hello")
is.character(NA)

# Obscure operators
21 %% 3
22 %% 3
21 %/% 3
22 %/% 3
```


## Imports

``` {r}
library(stats)
library(h2o)

Melanoma

library(MASS)

Melanoma

# Can also invoke package loading from the packages pane

# Can import functions in one line
dplyr::filter()

# Now's a good time to talk about finding help
# The help pane
?letters
??letters
```


## Methods

``` {r}
mean(c(1, 2, 3, 4))

base::mean(c(1, 2, 3, 4)) # Explain function function masking and that this alleviates, also means you don't have to load a whole library for one function

filter() #vs.
dplyr::filter()
```


## Statements

``` {r}
if (TRUE) {
  print("This worked!")
} else {
  print("Looks like it was false")
}

if (FALSE) {
  print("This worked!")
} else {
  print("Looks like it was false")
}

if (1 == 1) {
  print("This worked!")
} else {
  print("Looks like it was false")
}

if (2 != 2) {
  print("This worked!")
} else {
  print("Looks like it was false")
}

if (NA) {
  print("This worked!")
} else {
  print("Looks like it was false")
} # Breaks

ifelse(TRUE, 1, 0)
ifelse(FALSE, 1, 0)


numbers <- c(1 ,2 ,3)

for (num in numbers) {
  expression <- paste("num now equals", num)
  print(expression)
}

for (i in 1:3) {
  expression <- paste("num now equals", num)
  print(expression)
} # Breaks

for (i in 1:3) {
  expression <- paste("num now equals", numbers[i])
  print(expression)
}

# Nested fors
for (i in 1:3) {
  for (j in 1:3) {
    expression <- paste("i, j <- ", i , j)
    print(expression)
  }
}

# Vectorizing
print(numbers + 1)
```

## Syntax

``` {r}
# Numbers are always without quotes
1
2
3

# Strings are always in quotes
"hello"
"this is a string"
""

if 1==1
  print("It worked!")
  
""hello"

("hello"))
(("hello")
"hello"

{"hello"}
("hello")

# Vectors must be instantiated with c(...)
c(1,2,3)
c("hi", "hello", "I'm third")
1:50
c(1:50)

# Vectors of different types cause coercion
c(1, TRUE, "TRUE") # coerces all to string
c(1, 0, TRUE) # coerces to numeric
c(TRUE, FALSE, 1, 0) # coerces to numeric
as.logical(c(TRUE, FALSE, 1, 0)) # force coercion to logical

```

## Edge cases and errors

```{r}
Inf
1/0
-1/0
0/0
is.nan(Inf)
is.nan(-Inf)
is.na(Inf)
is.na(-Inf)
Inf == Inf
Inf * Inf
Inf + -Inf
Inf * -Inf

NaN == NA
is.na(NaN)
is.nan(NaN)
is.nan(NA)

NULL
is.na(NULL)

# Dates and times
```

## Subsetting

``` {r}
c(1,2,3,4,5)[c(1,2,3)]
c(1,2,3,4,5)[c(3,2,1)]
c(1,2,3,4,5)[c(2,3,1)]
c(1,2,3,4,5)[6]
```

## Intro to R and to Rmarkdown

``` {r}
Melanoma
View(Melanoma)
head(Melanoma)
head(Melanoma, n = 10)
tail(Melanoma)
str(Melanoma)
nrow(Melanoma)
ncol(Melanoma)
summary(Melanoma)

# Accessing
Melanoma$age
head(Melanoma$age)
summary(Melanoma$age)
```

